Engineering the Servo Web Browser Engine using Rust

### 2015-10-18 icse16-servo/abstract.tex
New code count numbers. Talk more about the soft. eng. bits

%!TEX root = paper.tex
%
\begin{abstract}
All modern web browsers --- Internet Explorer, Firefox, Chrome, Opera, and Safari ---
have a core rendering engine written in \Cplusplus{}.
This language choice was made because it affords the systems programmer complete control
of the underlying hardware features and memory in use, and it provides a transparent
compilation model.

@@

In this paper, we show how a language with an affine type system, regions, and many syntactic
features familiar to functional language programmers can be successfully used to build
state-of-the-art systems software.
We also outline several pitfalls encountered along the way and describe some potential
areas for future research.
\end{abstract}

>>>>

%!TEX root = paper.tex
%
\begin{abstract}
All modern web browsers --- Internet Explorer, Firefox, Chrome, Opera, and Safari ---
have a core rendering engine written in \Cplusplus{}.
This language choice was made because it affords the systems programmer complete control
of the underlying hardware features and memory in use, and it provides a transparent
compilation model.
Unfortunately, this language is both complex and highly susceptible to memory safety
issues that potentially lead to security holes.

@@

In this paper, we show how a language with an advanced type system can address many of
the most common security issues and software engineering challenges in other browser
engines, while still producing code that has the same performance and memory profile
required by these applications.
We also outline several pitfalls encountered along the way and describe some potential
areas for future research.
\end{abstract}

### 2015-10-18 icse16-servo/servo.tex
New code count numbers. Talk more about the soft. eng. bits

\subsection{Language interoperability}
Rust has nearly complete interoperability with C code, both exposing code to and using code from
C programs.
This support has allowed us to smoothly integrate with many browser libraries, which has been
critical for bootstrapping a browser without rewriting all of the lower-level libraries immediately, such as
graphics rendering code, the JavaScript engine, font selection code, etc.
\tblref{servo-loc} shows the breakdown between current lines of Rust code (including generated code that
handles interfacing with the JavaScript engine) and C code.
This table also includes test code, though the majority of that code is in HTML and JavaScript.
\begin{table}
  \begin{center}
    \begin{tabular}{r || r}
      Language & Lines of Code \\
      \hline
      C or \Cplusplus{} & 1,678,427 \\
      Rust & 410,817 \\
      HTML or JavaScript & 217,827 \\
    \end{tabular}%
  \end{center}%
  \caption{Lines of code in Servo}
  \label{servo-loc}
\end{table}

>>>>

\subsection{Language interoperability}
Rust has nearly complete interoperability with C code, both exposing code to and using code from
C programs.
The ability to use C code has allowed us to smoothly integrate with many browser libraries, which has been
critical for bootstrapping a browser without rewriting all of the lower-level libraries immediately, such as
graphics rendering code, the JavaScript engine, font selection code, etc.
Additionally, interoperation with C is required for components of a browser engine such as media decoders
for DRM content, which are sometimes delivered by vendors as a binary along with a C API.
\tblref{servo-loc} shows the breakdown between current lines of Rust code (including generated code that
handles interfacing with the JavaScript engine) and C code.
This table also includes test code, though the majority of that code is in HTML and JavaScript.
\begin{table}
  \begin{center}
    \begin{tabular}{r || r}
      Language & Lines of Code \\
      \hline
      C or \Cplusplus{} & 1,187,939 \\
      Rust & 816,158 \\
      HTML or JavaScript & 248,768 \\
    \end{tabular}%
  \end{center}%
  \caption{Lines of code in Servo (October 2015)}
  \label{servo-loc}
\end{table}

In the future, we hope to expose our Rust libraries written for Servo to C, in order to reuse them
in the Firefox web brower.
A first potential libraries is our Rust-based URL parser.\footnote{https://github.com/servo/rust-url}
While URL parsing seems like a simple and secure task, an audit of the public security bugs\footnote{Firefox public security bugs: http://mzl.la/1GkOiIC} shows that all four of the ones that come from URL parsing are related to either indexing out of
range of a raw memory buffer or using data that had already been freed --- both of which are
prevented by Rust code.
